Overlap between data sets and Volinia.
========================================================
`r as.character(Sys.time())`

```{r time, echo=FALSE}
starttime = Sys.time()
# copy and move an older version of the html file that will be produced
oldfn = "end_results.html"
if(file.exists(oldfn))
  {
    datestr = strsplit( as.character(file.info(oldfn)$ctime), " ")[[1]][1]
    backupname = sub(".html", paste("_",datestr, ".html", sep=""), oldfn)
    backupname = paste("not_in_github/", backupname, sep="")
    x=file.copy(oldfn,backupname, overwrite=TRUE )   
  }
```
<br/>
<br/>

We will check for

- concordance between our two approaches called **meta** and **merged**.
- concordance between our results and Volinia et al.
- evidence status of our findings based on a database of curated microRNAs.

<BR/>
<BR/>
```{r}
library(xtable)
if(!exists("inputisread"))
	source("read_input.r")
postfix = ".txt"
fdrCO = 0.05
```
## Meta vs Merged results.

Our data sets came from two providers, "AHUS" and "UCAM", using two different Agilent microRNA platforms. Batch effects were present as visualized in the QC report. We tested for differentially expressed genes using two approaches, combining all data in Limma using provider as a blocking factor ("merged" approach), and analyzing the two sets inpendetly with Limma and combine the p-values ("meta" approach).

Here we will list the agreemet between the two apporaches.

First a recap of the numer of samples and grouping based on annotation.

Sample count based on tissue type and provider.
```{r results='asis'}
print(xtable(table(sampleannotation[, c("provider" ,"tissue_type")]), 
             caption="", digits=3), 
      comment = TRUE,
      type = "html",
      html.table.attributes="CELLPADDING=5",
      include.rownames = TRUE)
```

Sample count based on PAM50 classification in each tissue and from each provider
```{r results='asis'}
for(p in unique(sampleannotation$provider))
{

	print(xtable(table(sampleannotation[sampleannotation$provider==p , c("tissue_type" ,"pam50_est")]), 
             caption=p, digits=3), 
      comment = TRUE,
      type = "html",
      html.table.attributes="CELLPADDING=5",
      include.rownames = TRUE)
}
```

Sample count based on IHC diagnosis in each tissue and from each provider
```{r results='asis'}
for(p in unique(sampleannotation$provider))
{

	print(xtable(table(sampleannotation[sampleannotation$provider==p , c("tissue_type" ,"IHC")]), 
             caption=p, digits=3), 
      comment = TRUE,
      type = "html",
      html.table.attributes="CELLPADDING=5",
      include.rownames = TRUE)
}
```




```{r}

#metafilefolder = "not_in_github/t-test"
metafilefolder = "not_in_github/lilianas"
mergedfilefolder = "output-analysis"
#metaresultfiles = list.files("lilianas/1000_permutations")
#mergedresultfiles = list.files("output-analysis-2014-08-05", pattern="difflist")

comparisons = c("DCIS-normal", "invasive-DCIS", "Normallike-DCIS", "LumA-DCIS", "LumB-DCIS", 
								"Her2-DCIS", "Basallike-DCIS")

# meta2merged = data.frame(metafn=character(0), mergedfn=character(0), direction=numeric(0), stringsAsFactors=FALSE)
# meta2merged["DCIS-normal",] = list("DCIS_to_normal_test", "difflist-DCIS-vs-normal.txt",-1)
# meta2merged["invasive-DCIS",] = list("DCIS_to_invasive_test", "difflist-invasive-vs-DCIS.txt",1)
# meta2merged["Normallike-DCIS",] = list("DCIS_to_normal(invasive)_test", "difflist-pam50-Normallike-vs-DCIS.txt",1)
# meta2merged["LumA-DCIS",] = list("DCIS_to_lumA_test", "difflist-pam50-LumA-vs-DCIS.txt",1)
# meta2merged["LumB-DCIS",] = list("DCIS_to_lumB_test", "difflist-pam50-LumB-vs-DCIS.txt",1)
# meta2merged["Her2-DCIS",] = list("DCIS_to_Her2_test", "difflist-pam50-Her2-vs-DCIS.txt",1)
# meta2merged["Basallike-DCIS",] = list("basal_to_DCIS_test", "difflist-pam50-Basallike-vs-DCIS.txt",-1)


meta2merged = data.frame(metafn=character(0), mergedfn=character(0), direction=numeric(0), stringsAsFactors=FALSE)
meta2merged["DCIS-normal",] = list("DCIS_to_normal_limma.txt", "difflist-DCIS-vs-normal.txt",-1)
meta2merged["invasive-DCIS",] = list("DCIS_to_invasive_limma.txt", "difflist-invasive-vs-DCIS.txt",1)
meta2merged["Normallike-DCIS",] = list("dcis_to_normal_like(invasive)_limma.txt", "difflist-pam50-Normallike-vs-DCIS.txt",1)
meta2merged["LumA-DCIS",] = list("dcis_to_lumA_limma.txt", "difflist-pam50-LumA-vs-DCIS.txt",1)
meta2merged["LumB-DCIS",] = list("dcis_to_lumB_limma.txt", "difflist-pam50-LumB-vs-DCIS.txt",1)
meta2merged["Her2-DCIS",] = list("dcis_to_her2_limma.txt", "difflist-pam50-Her2-vs-DCIS.txt",1)
meta2merged["Basallike-DCIS",] = list("basallike_to_dcis_limma.txt", "difflist-pam50-Basallike-vs-DCIS.txt",-1)

meta2merged["HER2neg_ERneg_PGRneg-DCIS",] = list("dcis_to_HER2neg_ERneg_PGRneg_limma.txt", "difflist-IHC-HER2neg_ERneg_PGRneg-vs-DCIS.txt",1)
meta2merged["HER2neg_ERpos-DCIS",] = list("dcis_to_HER2neg_ERpos_limma.txt", "difflist-IHC-HER2neg_ERpos-vs-DCIS.txt",1)
meta2merged["HER2pos_ERneg-DCIS",] = list("dcis_to_HER2pos_ERneg_limma.txt", "difflist-IHC-HER2pos_ERneg-vs-DCIS.txt",1)
meta2merged["HER2pos_ERpos-DCIS",] = list("dcis_to_HER2pos_ERpos_limma.txt", "difflist-IHC-HER2pos_ERpos-vs-DCIS.txt",1)


metatables = list()
mergedtables = list()
summarytab = data.frame( found_meta=integer(0),  found_merged=integer(0), 
                       found_both=integer(0),  stringsAsFactors=FALSE)

for(i in 1:nrow(meta2merged))
{   
    thiscomp = rownames(meta2merged)[i]
    fn = paste(metafilefolder, "/", meta2merged[thiscomp, "metafn"], sep="")
    if(file.info(fn)$size>10)
    	{
    thismetatab = read.table(fn, sep="\t", header=TRUE, stringsAsFactors=FALSE, fill=TRUE,
                                     strip.white=TRUE, comment.char ="")
    } else{
    	thismetatab = thismetatab[0,] # workaround to deal with empty files from meta.
    }
    fn = paste(mergedfilefolder, "/", meta2merged[thiscomp, "mergedfn"], sep="")
    thismergedtab = read.table(fn, sep="\t", header=TRUE, stringsAsFactors=FALSE, fill=TRUE,
                                     strip.white=TRUE, comment.char ="", row.names=1)
    
    #harmonizing
    colnames(thismetatab)[colnames(thismetatab)=="FC"] = "logFC"
    rownames(thismetatab) = thismetatab$MIMAT
    thismetatab$logFC = thismetatab$logFC * meta2merged[thiscomp, "direction"]
  #summarytab[thiscomp, "metafile"] = rownames(thisrows)[1]
  #summarytab[thiscomp, "mergedfile"] = thisrows[1, "mergedfn"]

  summarytab[thiscomp, "found_meta"] = nrow(thismetatab)
  b = thismergedtab$adj.P.Val <= 0.05
  summarytab[thiscomp, "found_merged"] = sum(b)
  
  commonMIMATID = intersect( rownames(thismetatab),  rownames(thismergedtab)[b])
  summarytab[thiscomp, "found_both"] = sum( thismetatab[commonMIMATID, "logFC"] * thismergedtab[commonMIMATID, "logFC"] > 0) 
  rm(commonMIMATID)
  metatables[[thiscomp]] = thismetatab
  mergedtables[[thiscomp]] = thismergedtab
}

```
<br/>
<br/>
<br/>
The number of microRNAs found differentially expressed between states, according to the two approaches. The direction of foldchange is taken into acount. 
```{r}
print(summarytab[,1:3])
```
For several of the comparisons, the shorter list seems to be mostly included in the longer.
<br/>
<br/>


## Our results vs. Volinia et al.´s results

```{r}
volinia_DCIS_vs_normal_file = "volinia-DCIS-vs-normal.txt"
volinia_invasive_vs_DCIS_file = "volinia-invasive-vs-DCIS.txt"
```

`r volinia_DCIS_vs_normal_file` and  `r volinia_invasive_vs_DCIS_file` are copy-pasted from Volinia et al.s [supplementary tables S1 and S2](http://www.pnas.org/content/suppl/2012/02/02/1200010109.DCSupplemental/pnas.201200010SI.pdf). The tables were reformatted from the pdf file into a more practical tabular text format. This is the results we aim to validate with the AHUS and UCAM data sets.

```{r}

voliniatab = list()
voliniatab[["invasive-DCIS"]] = read.table(file=paste(annotdir, "/", volinia_invasive_vs_DCIS_file, sep=""),
                                      sep="\t", check.names=TRUE, 
                                      stringsAsFactors=FALSE, header=TRUE)
voliniatab[["DCIS-normal"]]  = read.table(file=paste(annotdir, "/", volinia_DCIS_vs_normal_file, sep=""),
                                    sep="\t", check.names=TRUE, 
                                    stringsAsFactors=FALSE, header=TRUE)

for(n in names(voliniatab))
{
	# add a MIMATID 
	voliniatab[[n]] = as.data.frame(append(voliniatab[[n]], 
                  list(MIMATID=mimatmapping[paste("hsa-", voliniatab[[n]]$miRNA, sep="")]), 
                  after=1))
	# use log2 for foldchange in this document.
	voliniatab[[n]]$fc=log2(voliniatab[[n]]$fc)
	names(voliniatab[[n]])[names(voliniatab[[n]])=="fc"] = "log2FC"
	names(voliniatab[[n]])[names(voliniatab[[n]])=="pval"] = "FDR" # this was described as FDR in their paper.
}

```

```{r results='asis'}
print( xtable(head(voliniatab[[1]]), caption="Start of table S2 copied from Volinia et al.", digits=3), 
      comment = FALSE,
      type = "html",
      html.table.attributes="CELLPADDING=5",
      include.rownames = FALSE)
```


Compared to the original tables from Volinia, the fold change is here log2 transformed and "IDC" is renamed to "invasive". 


```{r results='asis'}

volinia_overlap = list()

for(contrast in names(voliniatab))
{
	#contrast = "invasive_vs_DCIS"
	overlaptab = voliniatab[[contrast]]
	overlaptab = overlaptab[!is.na(overlaptab$MIMATID),]
	rownames(overlaptab) = overlaptab$MIMATID
	overlaptab$Direction = ifelse(overlaptab$log2FC > 0, "UP", "DOWN")
	overlaptab = overlaptab[, c("miRNA", "MIMATID", "Direction","FDR")]
	names(overlaptab) = paste(names(overlaptab), "_vol", sep="")
		
	tab = mergedtables[[contrast]]
	a = intersect(  rownames(tab), rownames(overlaptab) )
	overlaptab$Direction_merged = NA
	overlaptab[a, "Direction_merged"] = ifelse(tab[a, "logFC"] > 0,  "UP", "DOWN")
	overlaptab$FDR_merged = NA
	overlaptab[a, "FDR_merged"] = tab[a, "adj.P.Val"]

	tab = metatables[[contrast]]
	a = intersect(  rownames(tab), rownames(overlaptab) )
	overlaptab$Direction_meta = NA
	overlaptab[a, "Direction_meta"] = ifelse(tab[a, "logFC"] > 0,  "UP", "DOWN")
	overlaptab$FDR_meta = NA
	overlaptab[a, "FDR_meta"] = fdrCO
	
	x = rowSums (overlaptab[,grepl("Direction", colnames(overlaptab))] == "UP")
	overlaptab$validated = NA
	overlaptab$validated[x %in% c(0,3)] = TRUE
	overlaptab$validated[x %in% c(1,2)] = FALSE
	overlaptab$validated = overlaptab$validated & (overlaptab$FDR_merged<=fdrCO & overlaptab$FDR_meta<=fdrCO)
	
# 	a = overlaptab$Direction_vol > 0 & overlaptab$Direction_merged > 0 & overlaptab$Direction_meta > 0
# 	b = overlaptab$Direction_vol < 0 & overlaptab$Direction_merged < 0 & overlaptab$Direction_meta < 0
# 	c = overlaptab$FDR_merged < 0.05 & overlaptab$FDR_meta < 0.05
# 	overlaptab$validated = (a|b) & c 		
	
	volinia_overlap[[contrast]] = overlaptab
}

```


```{r results='asis'}
x = order(volinia_overlap[["DCIS-normal"]]$FDR_merged)
print( xtable( volinia_overlap[["DCIS-normal"]][x,] , caption="microRNA found in Volinia compared to validation data", digits=3), 
      comment = FALSE,
      type = "html",
      html.table.attributes="CELLPADDING=5",
      include.rownames = FALSE)
```


## Checking results towards a curated list of microRNA´s
```{r}
	curatedfn = "curated_microRNA.csv"
  curatedtab  = read.table(paste("not_in_github", "/", curatedfn, sep=""), sep="\t", 
  												 header=TRUE, stringsAsFactors=FALSE, fill=TRUE,
                                     strip.white=TRUE, comment.char ="")
	curatedtab$miRNA = paste("hsa-", curatedtab$miRNA, sep="")
  # ad hoc fix
	curatedtab$miRNA = gsub("-mir-", "-miR-", curatedtab$miRNA)

	curatedtab$MIMAT = mimatmapping[curatedtab$miRNA]
	table(curatedtab$status)
	curatedtab = curatedtab[!is.na(curatedtab$MIMAT),]
	table(curatedtab$status)


```


Adding the curated status to the results

```{r results='asis'}

for(n in names(volinia_overlap))
{
	volinia_overlap[[n]]$curated = curatedtab[ match( rownames(volinia_overlap[[n]]), curatedtab$MIMAT ), "status"]
	voliniatab[[n]]$curated = curatedtab[ match( as.character(voliniatab[[n]]$MIMATID), curatedtab$MIMAT ), "status"]
}
x = order(volinia_overlap[["DCIS-normal"]]$FDR_merged)
print( xtable( volinia_overlap[["DCIS-normal"]][x,] , caption="With curated status", digits=3), 
      comment = FALSE,
      type = "html",
      html.table.attributes="CELLPADDING=5",
      include.rownames = FALSE)
```


The difference in curated status for the filtered out microRNAs and the ones kept as expressed, and the ones found as significantly differentially expressed between the "DCIS" and "normal" stat. AHUS only for the "expressed" and "nonexpressed". Both UCAM and AHUS in the significant. 


```{r results='asis'}

#	table(curatedtab$status[  curatedtab$MIMAT %in% rownames(mergedtables[[1]]) ])

expressed =	table(curatedtab$status[curatedtab$MIMAT %in% filteredAHUSMIMAT])
nonexpressed=	table(curatedtab$status[curatedtab$MIMAT %in% unfilteredAHUSMIMAT[!unfilteredAHUSMIMAT %in% filteredAHUSMIMAT]])

genelist = intersect( rownames(mergedtables[["DCIS-normal"]])[mergedtables[["DCIS-normal"]]$adj.P.Val<0.05], 
					 rownames(metatables[["DCIS-normal"]]) )
significant = table(curatedtab$status[curatedtab$MIMAT %in% genelist])


print( xtable( rbind(nonexpressed,expressed,significant), caption="", digits=3), 
      comment = FALSE,
      type = "html",
      html.table.attributes="CELLPADDING=5",
      include.rownames = TRUE)


```

